The Effect of Fictional Reappraisal on Subjective Ratings Toward Images

### 2025-09-02 paper/manuscript.qmd
started intro

>>>>

<!-- AI and Reality distinction
  - rise of AI
  - societal consequences
  - indistinguishable nature of artificial vs real content -->

Recent advances in artificial intelligence (AI) have introduced new challenges for human cognition, particularly regarding the ability to distinguish between authentic and fabricated experiences [@miller2023ai]. 
The stakes of such perceptual uncertainty are considerable, with misinformation representing one of the most pressing societal consequences [@kreps2022all].
For instance, deepfakes (face-swapping technologies that enable the creation of realistic fake images and videos) have already been employed to fabricate convincing political speeches that appear genuine [@meskys2020regulating].
When deployed in electoral contexts, such content has the potential to distort public opinion and undermine democratic processes by rendering the truth unclear [@graber2021artificial]. 
Yet political disinformation is only one domain in which synthetic media exerts influence. 
The increased accessibility of generative technologies means that manipulated or entirely artificial content now permeates social media, entertainment, immersive environments, and interpersonal communication [@nightingale2022ai].  
In these settings, the boundaries between the “real” and the “artificial” are becoming progressively less discernible, raising fundamental questions about how individuals perceive and evaluate reality.

<!-- Reality Perception
  - ambiguous stimuli
  - perceptual cues, emotional influence, reality monitoring
  - AI getting better at producing realistic content -->

A key challenge lies in the prevalence of ambiguous stimuli, that is images, texts, videos, or environments whose authenticity cannot be easily established. 
While artificial stimuli once carried perceptual markers that betrayed their inauthenticity, such as distortions in early computer-generated imagery [@corvi2023detection; @mcdonnell2010face], these limitations are rapidly disappearing. 
In the domain of face generation especially, current algorithms now produce synthetic images that are virtually indistinguishable from genuine photographs [@moshel2022you; @nightingale2022ai]. 
As these technologies approach perfect realism, the study of reality perception becomes not only a theoretical concern but also a practical imperative, with direct implications for information security, social trust, and the ethical deployment of AI.

<!-- Affective Responses and Reality Judgments
  - Link perception of realism with emotional engagement: arousal, valence, enticement.
  - Mention theories such as Affective Reality Theory (Makowski, 2023) if relevant, or more general emotion–cognition links.
  - Emotions may not just follow perception, but also shape how we judge what is real. -->

<!-- Sexual Stimuli and Pornography Consumption
  - Motivation behind inclusion of sexual/non-sexual content (e.g., sexual images are affectively intense, highly relevant for arousal/enticement measures).
  - Bring in literature on pornography consumption and attitudes toward AI in intimate contexts (questionnaires)
  - Highlight that this domain is both theoretically interesting (strong affective reactions) and socially relevant (AI-generated sexual content already exists, raising ethical debates). -->

### 2025-09-02 paper/manuscript.tex
started intro

\par{     \begin{tcolorbox}[enhanced jigsaw, arc=.35mm, opacityback=0, colback=white, breakable, left=2mm, leftrule=.75mm, rightrule=.15mm, colframe=quarto-callout-note-color-frame, toprule=.15mm, bottomrule=.15mm]

@@

@@

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-de2015jspsych}
De Leeuw, J. R. (2015). jsPsych: A JavaScript library for creating
behavioral experiments in a web browser. \emph{Behavior Research
Methods}, \emph{47}(1), 1--12.

@@

@@

>>>>

\par{     \begin{tcolorbox}[enhanced jigsaw, bottomrule=.15mm, colframe=quarto-callout-note-color-frame, left=2mm, breakable, colback=white, opacityback=0, rightrule=.15mm, toprule=.15mm, leftrule=.75mm, arc=.35mm]

@@

Recent advances in artificial intelligence (AI) have introduced new
challenges for human cognition, particularly regarding the ability to
distinguish between authentic and fabricated experiences
(\citeproc{ref-miller2023ai}{Miller et al., 2023}). The stakes of such
perceptual uncertainty are considerable, with misinformation
representing one of the most pressing societal consequences
(\citeproc{ref-kreps2022all}{Kreps et al., 2022}). For instance,
deepfakes (face-swapping technologies that enable the creation of
realistic fake images and videos) have already been employed to
fabricate convincing political speeches that appear genuine
(\citeproc{ref-meskys2020regulating}{Meskys et al., 2020}). When
deployed in electoral contexts, such content has the potential to
distort public opinion and undermine democratic processes by rendering
the truth unclear (\citeproc{ref-graber2021artificial}{Graber-Mitchell,
2021}). Yet political disinformation is only one domain in which
synthetic media exerts influence. The increased accessibility of
generative technologies means that manipulated or entirely artificial
content now permeates social media, entertainment, immersive
environments, and interpersonal communication
(\citeproc{ref-nightingale2022ai}{Nightingale \& Farid, 2022}).\\
In these settings, the boundaries between the ``real'' and the
``artificial'' are becoming progressively less discernible, raising
fundamental questions about how individuals perceive and evaluate
reality.

A key challenge lies in the prevalence of ambiguous stimuli, that is
images, texts, videos, or environments whose authenticity cannot be
easily established. While artificial stimuli once carried perceptual
markers that betrayed their inauthenticity, such as distortions in early
computer-generated imagery (\citeproc{ref-corvi2023detection}{Corvi et
al., 2023}; \citeproc{ref-mcdonnell2010face}{McDonnell \& Breidt,
2010}), these limitations are rapidly disappearing. In the domain of
face generation especially, current algorithms now produce synthetic
images that are virtually indistinguishable from genuine photographs
(\citeproc{ref-nightingale2022ai}{Nightingale \& Farid, 2022};
\citeproc{ref-moshel2022you}{\textbf{moshel2022you?}}). As these
technologies approach perfect realism, the study of reality perception
becomes not only a theoretical concern but also a practical imperative,
with direct implications for information security, social trust, and the
ethical deployment of AI.

@@

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-corvi2023detection}
Corvi, R., Cozzolino, D., Zingarini, G., Poggi, G., Nagano, K., \&
Verdoliva, L. (2023). On the detection of synthetic images generated by
diffusion models. \emph{ICASSP 2023-2023 IEEE International Conference
on Acoustics, Speech and Signal Processing (ICASSP)}, 1--5.

\bibitem[\citeproctext]{ref-de2015jspsych}
De Leeuw, J. R. (2015). jsPsych: A JavaScript library for creating
behavioral experiments in a web browser. \emph{Behavior Research
Methods}, \emph{47}(1), 1--12.

\bibitem[\citeproctext]{ref-graber2021artificial}
Graber-Mitchell, N. (2021). \emph{Artificial illusions: Deepfakes as
speech}.

@@

\bibitem[\citeproctext]{ref-kreps2022all}
Kreps, S., McCain, R. M., \& Brundage, M. (2022). All the news that's
fit to fabricate: AI-generated text as a tool of media misinformation.
\emph{Journal of Experimental Political Science}, \emph{9}(1), 104--117.

@@

\bibitem[\citeproctext]{ref-mcdonnell2010face}
McDonnell, R., \& Breidt, M. (2010). Face reality: Investigating the
uncanny valley for virtual faces. In \emph{ACM SIGGRAPH ASIA 2010
sketches} (pp. 1--2).

\bibitem[\citeproctext]{ref-meskys2020regulating}
Meskys, E., Kalpokiene, J., Jurcys, P., \& Liaudanskas, A. (2020).
Regulating deep fakes: Legal and ethical considerations. \emph{Journal
of Intellectual Property Law \& Practice}, \emph{15}(1), 24--31.

\bibitem[\citeproctext]{ref-miller2023ai}
Miller, E. J., Steward, B. A., Witkower, Z., Sutherland, C. A.,
Krumhuber, E. G., \& Dawel, A. (2023). AI hyperrealism: Why AI faces are
perceived as more real than human ones. \emph{Psychological Science},
\emph{34}(12), 1390--1403.

\bibitem[\citeproctext]{ref-nightingale2022ai}
Nightingale, S. J., \& Farid, H. (2022). AI-synthesized faces are
indistinguishable from real faces and more trustworthy.
\emph{Proceedings of the National Academy of Sciences}, \emph{119}(8),
e2120481119.
